Viscous Erosion in Porous Media

### 2018-03-26 docs/methods/erosion_methods.tex
NM

These fixed-area simulations enable us to precisely characterize the changing geometry as it approaches a terminal state. In particular, we measure the aspect ratio of the body (the width to height) and the opening angle of the front/rear corner as they vary in time. In these tests, we keep the spatial resolution fixed, $N_\iin = 1024$, and show three different cases of the smoothing parameters: $\eps = \sigma = 20/1024, 10/1024,$ and 5/1024.

The fixed-area simulations allow for us to precisely measure characteristics of the shape. In particular, Fig.~\ref{fig:arangle} shows measurements of the body's aspect ratio (its width to height) and the front opening angle as they change in time. We show these measurements for three simulations in which the smoothing parameters, $\eps$ and $\sigma$, are sequentially reduced, allowing us to assess convergence in these parameters. As shown in Fig.~\ref{fig:arangle}(a), the aspect ratio begins at one for the initial circular geometry and increases in time as erosion sculpts a more slender shape. The aspect ratio reaches a long time limit of about 2.7 in the most resolved simulation. Fig.~\ref{fig:arangle}b shows the body's opening angle, which we measure by fitting the tangent angle $\theta(\alpha)$ with a 7th degree polynomial over a region excluding the front/rear stagnation point (where a near-corner is approximated by a smooth curve of very high curvature). The polynomial fit is then extrapolated to the stagnation point to obtain an angle estimate (subject to some uncertainty, which we estimate below). As seen in Fig.~\ref{fig:arangle}a, the opening angle is initially $180^\circ$, corresponding to the initially circular geometry. Erosion causes the angle to form after a short time, and the angle becomes increasingly sharp as it converges to the value $102^\circ$ in the most resolved simulation. This coincides exactly with the valued predicted by our local solutions.

>>>>

These fixed-area simulations enable us to precisely characterize the changing geometry as it approaches a terminal state. In particular, we measure the aspect ratio of the body (the width to height) and the opening angle of the front/rear corners. In these tests, we keep the spatial resolution fixed, $N_\iin = 1024$, and show three different cases of the smoothing parameters ($\eps = \sigma = 20/1024, 10/1024,$ and 5/1024) so that we can assess how the numerical smoothing influences shape. Figure\ref{fig:arangle}(a) shows the aspect ratio as it varies in time. The initially circular geometry has an aspect ratio of unity, and, as erosion carves a more slender form, the aspect ratio increases monotonically. The three simulations are all very close to one another, which demonstrates convergence as the smoothing parameters are reduced. In the most accurate simulation (i.e.~the least smoothing), the aspect ratio tends to a value of 2.7 in long time.

Next, we aim to measure the opening angle that develops at the front and rear of the body. Since these simulations are actually smooth on a fine scale, a true corner does not exist and we must instead estimate the apparent angle that results from a small region of high curvature. To this end, we first fit the tangent angle $\theta(\alpha)$ with a 7th-degree polynomial over a region that excludes a small vicinity of the front/rear stagnation points ($\alpha = 0, 0.5$). We then extrapolate the polynomial to the stagnation points which gives an estimate of the angle there. This protocol is subject to some uncertainty, which we will examine further below.

As seen in Fig.~\ref{fig:arangle}a, the opening angle is initially $180^\circ$, corresponding to the initially circular geometry. Erosion causes the angle to form after a short time, and the angle becomes increasingly sharp as it converges to the value $102^\circ$ in the most resolved simulation. This coincides exactly with the valued predicted by our local solutions.
